### 2015-04-24 article/article.tex
Web of Science keyword search and article skeleton

\documentclass[12pt,letterpaper]{article} % for a short document

% Bibliography
\usepackage[authordate,strict,autolang=hyphen,bibencoding=inputenc,doi=true,isbn=false,annotation=false]{biblatex-chicago}
\addbibresource{../references/references.bib}

\title{Reproducible Methods for Urban Form Analysis: A Case Study of PostGIS}
\author{Phil Hurvitz, \ldots}
\date{}

%%% BEGIN DOCUMENT

\begin{document}

\maketitle

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris in
hendrerit ligula. Cras malesuada auctor egestas. In eu placerat
dolor. Vestibulum rhoncus risus nulla. Praesent consequat diam nec
neque hendrerit, ac convallis turpis feugiat. Nulla quis nisi
lectus. Suspendisse ipsum eros, facilisis at nulla sed, dignissim
rhoncus erat. Suspendisse ut odio quis justo egestas condimentum eu
vitae mauris. Donec tincidunt sapien ex, at porta arcu scelerisque
in. Nunc porttitor risus et rhoncus venenatis. Sed aliquet rhoncus
interdum. Quisque nec lacinia risus. Nulla mi nulla, varius ac elit
iaculis, blandit lobortis augue. In pharetra erat ut ipsum pulvinar,
id blandit purus mollis.

\cite{Forsyth2014sausage}

\printbibliography

\end{document}

### 2015-05-20 analysis/000-data-aquisition.sh
add ETL scripts, start of Makefile, initial SB run

@@ -1,0 +1,3 @@
+#!/usr/bin/env bash
+
+# Nothing to do; handled by Makefile at top

### 2015-05-20 analysis/010-db-create.sh

@@ -1,0 +1,24 @@
+#!/usr/bin/env bash
+
+# Assumption here is that PostGIS is running locally, that the
+# current user has administrative rights to the DB, and that the
+# postgres user binaries are all in the current user's PATH.
+#
+# WARNING: If you happen to have a db named pgismethods, well, you'll
+# have a different one after running this script.
+
+DB=pgismethods
+
+# Drop the DB
+dropdb $DB
+
+# Create the DB
+createdb $DB
+
+# Enable PostGIS extensions
+echo "CREATE EXTENSION postgis; CREATE EXTENSION postgis_topology;" |
+    psql $DB
+
+# Enable PgRouting extension
+echo "CREATE EXTENSION pgrouting;" |
+    psql $DB

### 2015-05-20 analysis/020-db-loaddata.sh

@@ -1,0 +1,27 @@
+#!/usr/bin/env bash
+
+DB=pgismethods
+
+# Use osm2pgrouting to build the table
+osm2pgrouting -file "data/net/washington-latest.osm" \
+              -conf "analysis/020-mapconfig.xml" \
+              -dbname $DB \
+              -user `whoami` \
+              -clean
+
+
+# Convert / project the network data to a WA-North NAD83 HARN ft
+echo "SELECT addgeometrycolumn('ways', 'geom_2926',
+      2926, 'LINESTRING', 2);
+      UPDATE ways SET geom_2926 = st_transform(the_geom, 2926);
+      CREATE INDEX ways_geom_2926_gidx ON ways
+      USING gist(geom_2926);" |
+    psql $DB
+
+# Ditto the vertices
+echo "SELECT addgeometrycolumn('ways_vertices_pgr',
+      'geom_2926',2926,'POINT',2);
+      UPDATE ways_vertices_pgr set geom_2926 = st_transform(the_geom, 2926);
+      CREATE INDEX ways_vertices_pgr_geom_2926_gidx ON ways
+      USING gist(geom_2926);" |
+    psql $DB

### 2015-05-20 analysis/025-db-loadfunctions.sh

@@ -1,0 +1,6 @@
+#!/usr/bin/env bash
+DB=pgismethods
+
+cat ./025-sausagebuffer.sql |
+    psql $DB
+

### 2015-05-20 article/article.tex
Initial article intro

@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris in
-hendrerit ligula. Cras malesuada auctor egestas. In eu placerat
-dolor. Vestibulum rhoncus risus nulla. Praesent consequat diam nec
-neque hendrerit, ac convallis turpis feugiat. Nulla quis nisi
-lectus. Suspendisse ipsum eros, facilisis at nulla sed, dignissim
-rhoncus erat. Suspendisse ut odio quis justo egestas condimentum eu
-vitae mauris. Donec tincidunt sapien ex, at porta arcu scelerisque
-in. Nunc porttitor risus et rhoncus venenatis. Sed aliquet rhoncus
-interdum. Quisque nec lacinia risus. Nulla mi nulla, varius ac elit
-iaculis, blandit lobortis augue. In pharetra erat ut ipsum pulvinar,
-id blandit purus mollis.
+\section*{Background}
+When delineating travel catchment areas or travel sheds, network-based
+buffers are often preferable to circular, great arc buffers in that
+network buffers more realistically reflect the distance limitations
+imposed by different street patterns. \textcite{Forsyth2014sausage}
+identified a threat to replicability for studies using such buffers,
+due to changes over the years in how ESRI's market-leading GIS
+software has calculated such service areas. In response,
+\citeauthor{Forsyth2014sausage} proposed ``sausage'' network buffers,
+created via a transparent algorithm, for use in measuring aspects of
+urban form related to food and physical activity. Presented as
+replicable and cross-platform, the implementation described in that
+article nevertheless relies on ESRI's commercial GIS and Network
+Analyst extensions, and is thus tied to the Microsoft Windows
+platform. This paper validates the the sausage buffer technique with
+particular attention paid to reproducibility and cross-platform
+implementation using readily available free and open source
+software. Then, because buffer creation is often an early step in a
+larger analysis, we present a case study demonstrating the
+complementarity of our sausage buffering implementation with a typical
+follow-on analysis within the PostGIS platform.
 
-\cite{Forsyth2014sausage}
+% Let's mention something about reproducibility up front.
+The methods presented here are in part a response to a growing call
+from across numerous disciplines to ensure computer aided analyses are
+reproducible. The basic standard for reproducibility is that the data
+used are available, and that analyses are shared in the form of
+computer code \cite{Peng2011computational}. In that sense, the methods
+presented here are reproducible because analyses are executable code
+rather than a set of manual procedures, and that those pieces of code
+and the data are publicly available. Further, building these methods
+atop free and open source software achieves a higher standard of
+reproducibility by eliminating dependencies on any ``black box''
+computer code. In principle, this means that every piece of software
+used in these methods is open to scrutiny.
+
+% Relevance to planning and epidemiology
+Developing geoprocessing methods around free and open source tools
+holds particular benefits for disciplines such as epidemiology, public
+policy, and urban planning because of the reduced barriers of adoption
+for practitioners and advocates outside of academia. Academic
+institutions and larger public agencies may be able to afford software
+licenses. However, smaller agencies or advocacy groups wishing to
+perform, for example, walkshed analysis might otherwise be unable to
+make use of analyses reliant on proprietary software.
+
+% Not a literature review, per se. Since we are developing a methods
+% paper largely in response to one particularly important method, I'm
+% not sure how important a full-blown review is. Open to suggestions,
+% however. This could potentially be split off and developed into
+% something more complete. 
+Finally, the PostGIS platform is especially beneficial in that it
+naturally facilitates reproducible scripted analyses, as well as can
+provide the core infrastructure for data management on a research
+project. We are not the first to identify PostGIS for use in
+research. In a keyword search for ``PostGIS'' on ISI Web of Science,
+we identified 35 articles across a number of disciplines describing
+their use of PostGIS. In the transportation sector,
+\textcite{Wang2015routable} used PostGIS to store and analyze a large
+number of GPS traces in order to infer characteristics about roadways
+such as position and rules, for the purposes of generating routable
+roadway networks. In another recent article,
+\textcite{Brovelli2015FOSS} describe the use of PostGIS together with
+GeoServer to collect data about roadway pavement conditions from the
+general public. Notable in this study is the potential for PostGIS in
+supporting digitally-mediated public participation.
+
+\section*{Methods}
+Using the process described by \textcite{Forsyth2014sausage}, we
+implemented the sausage buffering technique using the PostgreSQL
+relational database with PostGIS and PgRouting extensions. We then
+implemented the same technique in ArcGIS Desktop X.XX, as a basis for
+comparison with the results presented by
+\citeauthor{Forsyth2014sausage}.
+
+Using the two buffering implementations, we generated two sets of
+sausage buffers around a common set of N points. These points were
+randomly selected from non-water areas of <SOME GEOGRAPHY, probably
+King County>. We then make several cross-comparisons between the two
+sets of buffers, matching buffers according to a unique identifier
+assigned to the initial set of points. First we compare the difference
+in area between each set of buffers. Then, to assess the possibility
+of spatial misalignment between similarly sized buffers, we computed
+the symmetric difference of the two sets of sausage buffers and then
+calculated the area of the non-overlapping regions.
+
+\section*{Results}
+
+\section*{Discussion}
+
+\section*{Conclusions}
 

### 2015-06-15 article/article.tex
first draft uber roll-up

--- 2015-04-24 article/article.tex
@@
-\documentclass[12pt,letterpaper]{article} % for a short document
+\documentclass[11pt,letterpaper]{article} % for a short document
 
@@
 \usepackage[authordate,strict,autolang=hyphen,bibencoding=inputenc,doi=true,isbn=false,annotation=false]{biblatex-chicago}
+\usepackage{graphicx}
+\DeclareGraphicsExtensions{.pdf,.png,.jpg}
+\usepackage[section]{placeins}
 \addbibresource{../references/references.bib}
+\usepackage{booktabs}
 
-\title{Reproducible Methods for Urban Form Analysis: A Case Study of PostGIS}
+\title{Reproducible Sausage Buffers for Urban Form Analysis: A Case Study of PostGIS}
 \author{Phil Hurvitz, \ldots}
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris in
-hendrerit ligula. Cras malesuada auctor egestas. In eu placerat
-dolor. Vestibulum rhoncus risus nulla. Praesent consequat diam nec
-neque hendrerit, ac convallis turpis feugiat. Nulla quis nisi
-lectus. Suspendisse ipsum eros, facilisis at nulla sed, dignissim
-rhoncus erat. Suspendisse ut odio quis justo egestas condimentum eu
-vitae mauris. Donec tincidunt sapien ex, at porta arcu scelerisque
-in. Nunc porttitor risus et rhoncus venenatis. Sed aliquet rhoncus
-interdum. Quisque nec lacinia risus. Nulla mi nulla, varius ac elit
-iaculis, blandit lobortis augue. In pharetra erat ut ipsum pulvinar,
-id blandit purus mollis.
+\section*{Background}
+When delineating travel catchment areas or travel sheds, network-based
+buffers are often preferable to circular, great arc buffers in that
+network buffers more realistically reflect the distance limitations
+imposed by different street patterns. \textcite{Forsyth2014sausage}
+identified a threat to replicability for studies using such
+network-based buffers, due to changes over the years in how ESRI's
+market-leading GIS software has calculated them. In response,
+\citeauthor{Forsyth2014sausage} proposed ``sausage'' network buffers,
+created via a transparent algorithm, for use in measuring aspects of
+urban form related to food and physical activity. Presented as
+replicable and cross-platform, the implementation described in that
+article nevertheless relies on ESRI's commercial GIS and Network
+Analyst extensions, and is thus tied to the Microsoft Windows
+platform. Their paper called on others to validate sausage buffers
+created via other GIS software on other platforms. This paper answers
+that call, validating the the sausage buffer technique with particular
+attention paid to reproducibility and cross-platform implementation
+using readily available free and open source software.
 
-\cite{Forsyth2014sausage}
+% Let's mention something about reproducibility up front.
+The methods presented here are also in part a response to a growing
+call from across numerous disciplines to ensure computer aided
+analyses are reproducible. The basic standard for reproducibility is
+that the data used are available, and that analyses are shared in the
+form of computer code \cite{Peng2011computational}. In that sense, the
+methods presented here are reproducible because analyses are fully
+automated using computer code, rather than being reliant on a set of
+manual procedures. Those pieces of code and the data are publicly
+available.\footnote{need to package up code and data, upload to UW's
+  repository, and then see about generating a DOI.}  Further, building
+these methods atop free and open source software achieves a higher
+standard of reproducibility by eliminating dependencies on any ``black
+box'' computer code. In principle, this means that every piece of
+software used in these methods is open to scrutiny.
+
+% Relevance to planning and epidemiology
+Developing geoprocessing methods around free and open source tools
+holds particular benefits for disciplines such as epidemiology, public
+policy, and urban planning because of the reduced barriers of adoption
+for practitioners and advocates outside of academia. Academic
+institutions and larger public agencies may be able to afford software
+licenses. However, smaller agencies or advocacy groups wishing to
+perform this type of walkshed analysis might otherwise be unable to
+make use of analyses reliant on proprietary software simply due to the
+cost of software licenses.
+
+% Not a literature review, per se. Since we are developing a methods
+% paper largely in response to one particularly important method, I'm
+% not sure how important a full-blown review is. Open to suggestions,
+% however. This could potentially be split off and developed into
+% something more complete. 
+Finally, the PostGIS platform is especially beneficial in that it
+naturally facilitates reproducible, scripted analyses, as well as can
+provide the core infrastructure for data management on a research
+project. \textcite{hurvitz2014emerging} describe the use of
+PostgreSQL/PostGIS to manage and process geospatial and timeseries
+data from a variety of sources including GPS and
+accelerometers. However, we are neither first nor alone in identifying
+PostGIS for use in research. In a keyword search for ``PostGIS'' using
+ISI Web of Science, we identified 35 articles across a number of
+disciplines describing the use of PostGIS in research. In the
+transportation sector, \textcite{Wang2015routable} also used PostGIS
+to store and analyze a large number of GPS traces in order to infer
+characteristics about roadways such as position and rules, for the
+purposes of generating routable roadway networks. In another recent
+article, \textcite{Brovelli2015FOSS} describe the use of PostGIS
+together with GeoServer to collect data about roadway pavement
+conditions from the general public. Notable in this study is the
+potential for PostGIS in supporting digitally-mediated public
+participation.
+
+\section*{Methods}
+% Set-up and software PgRouting 2.x on OpenBSD 5.7 i5
+In this section we describe how, following the process described by
+\textcite{Forsyth2014sausage, Forsyth2012proto}, we implemented the
+sausage buffering technique using the PostgreSQL relational database
+with PostGIS and PgRouting extensions. Using that technique we
+generated 200 such buffers around randomly selected points. We then
+compare those results to buffers analogously created in ArcGIS Desktop
+10.2.X, for use as a basis for comparison with the results presented
+by \citeauthor{Forsyth2014sausage}. Table \ref{tab:software} lists
+software versions used in this analysis.
+
+\begin{table}[h!]
+  \centering
+  \caption{Software used to generate PostGIS and ESRI sausage buffers}
+  \begin{tabular}{llr}
+  \toprule
+  Buffer type & Software & Version \\
+  \midrule
+PostGIS/PgRouting & OpenBSD (Operating System) & 5.7 \\
+              & PostgreSQL & 9.4.2 \\
+              & PostGIS Extensions & 2.1.7 \\
+              & PgRouting & 2.0.0 \\
+              & Osm2pgrouting & 2.0.0 \\
+              &  &  \\
+  ESRI & Microsoft Windows & Eric? \\
+              & Arc(Map)? & 10.2.X Eric? \\
+  \bottomrule
+\end{tabular}
+\label{tab:software}
+\end{table}
+
+In overview, the process for creating a sausage buffer entails three
+key steps. First, a point of origin is identified on a network of
+transportation facilities. Next, a distance threshold is selected. The
+network is traversed outward along all possible paths from the point
+of origin until reaching the distance threshold. The resulting subset
+reflects portions of the network reachable within the chosen
+distance. Finally, the segments of the network subset are buffered
+according to some chosen buffer radius, resulting in a set of polygons
+reminiscent of sausages. These are combined into a single polygon,
+forming the final buffer. Figure \ref{fig:sausage-steps} illustrates
+these steps.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\textwidth]{./figs/sausage-steps}
+  \caption{From left to right, the three primary steps in creating a
+    sausage buffer. First, an origin is selected. Next, the portions
+    of the network reachable within a given distance threshold (here,
+    1km) are identified. Finally, lines are combined and buffered.}
+  \label{fig:sausage-steps}
+\end{figure}
+
+%\subsection*{Data sources and network import}
+To create the routable network required for generating the buffers, we
+imported the Washington State portion of the Open Street Map (OSM)
+into PostGIS / PgRouting using the open source osm2pgrouting
+utility. Because much of our work deals with pedestrian travel, we
+excluded highway facilities, which often do not allow for pedestrian
+access. The osm2pgrouting import strategy yielded two primary
+benefits. First, osm2pgrouting facilitates reproducible data
+importation through the use of an XML-based configuration file. This
+allows facility exclusions, such as our exclusion of highways for
+pedestrians, to be carried out consistently. Secondly, this strategy
+preserves the topology of the OSM network, rather than inferring a
+topology from the network's geometry. This second point has
+implications for comparability of buffers that we will discuss later.
+
+%\subsection*{Generation of buffers}
+
+% Creation of buffer code in database
+% Random sample of 200 vertices as points to construct buffers around
+% Define sausage buffering procedure in PLSQL
+% Construct 1000m buffer (net dist), 100m (arc dist)
+% Rounded versus squared ends
+We then implemented the sausage buffering algorithm in the PL/PgSQL
+programming language, which is an extended variant of the SQL
+programming language. This allowed us to define the sausage buffering
+procedures and then store them together with data in the
+database. Here, we defined two such functions. The first calculates a
+single sausage buffer around a single point of origin. The second uses
+the first function to calculate sausage buffers around an arbitrary
+number of points.
+
+Because a point of origin may not be located precisely on the network,
+the sausage buffering function begins by identifying the nearest
+vertex in the network to the identified point of origin. To limit the
+required number of distance calculations required, the function then
+identifies vertices within a circular buffer with radius equal to the
+chosen network distance threshold. This buffer represents the maximum
+possible distance one could travel on the network in any given
+direction. We then calculate the distance to each vertex within this
+circular buffer, discarding routes exceeding the distance threshold,
+and retaining those less than or equal to the threshold.\footnote{We
+  believe this approach to be computationally inefficient, however
+  necessary due to the lack of a built-in service area function in
+  PgRouting. Our hope is that future versions of PgRouting will
+  include service area analysis, thus simplifying and speeding up
+  creation of these buffers.} The resulting set of lines are then
+buffered according to the buffer radius, and the resulting polygons
+are dissolved into a single geometry.
+
+After implementing the sausage buffering technique in PL/PgSQL, we
+then used it to generate buffers originating from 200 randomly
+selected points. To ensure that these points were not located in
+roadless areas of the state, we selected these points from the
+vertices in the roadway network. Then, in order to compare buffers
+created using our implementation with buffers described in the
+original \citeauthor{Forsyth2014sausage} paper, we created buffers
+around the same set of 200 points using the protocols described in
+\textcite{Forsyth2012proto}, with one change. In that document, the
+protocols call for ``flat'' ended buffers, however in the more
+recently published article, \citeauthor{Forsyth2014sausage} describe
+rounded buffer ends. We chose to implement the rounded buffer ends as
+we believe they are more theoretically defensible.
+
+We carried out this process of duplication manually as described in
+that document using ArcGIS 10.X. In addition, to ensure a maximum of
+consistency between the datasets used to generate buffers, we exported
+both the random points as well as the road network. We exported both
+of these datasets as shapefiles, which were then imported into
+ArcGIS. Worth noting is that, as described in the
+\textcite{Forsyth2012proto} protocols and replicated by us, the
+roadway network was imported without preserving the original OSM
+topology. While we believe the resulting networks to be largely the
+same, we note it as a potential source of difference in the two
+analyses.
+
+Finally, we made two cross-comparisons between the two sets of
+buffers. Here we matched pairs of buffers according to a unique
+identifier assigned to the initial set of points. First we compared
+the areas of the two sets of buffers. To assess whether differences in
+the areas between these two groups were statistically significant, we
+used a Bayesian procedure somewhat analogous to a t-test called BEST
+\autocite{Kruschke2014,Kruschke2013}. Second, to assess the possibility of
+spatial misalignment between similarly sized buffers, we computed the
+symmetric difference of the two sets of sausage buffers and then
+calculated the area of the non-overlapping regions. The results of
+these comparisons are presented in the following section.
+
+\section*{Results and Discussion}
+On comparing the areas of the two sets of sausage buffers, we found
+that, overall, they were very similar, with perhaps more
+PostGIS-generated buffers at slightly smaller sizes. Figure
+\ref{fig:area_dens} shows the distribution of these two sets of
+buffers. Both exhibit the same somewhat bimodal shape, with greater
+differences observed among smaller buffers.
+
+That apparent difference in size did not, however, constitute a
+statistically significant difference between the two sets of
+buffers. We estimated the mean difference's 95\% confidence interval
+as ranging from -.23 to 0.04. Because a mean difference of zero falls
+within our 95\% confidence interval, we cannot say for certain that
+PostGIS generated sausage buffers will generally be smaller than
+buffers created in ArcGIS.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.65\textwidth]{../results/area_dens}
+  \caption{Density plot showing the distributions PostGIS- and
+    ESRI-generated sausage buffers across the range of sizes.}
+  \label{fig:area_dens}
+\end{figure}
+
+When we compare the symmetric differences between the two sets of
+buffers we see (figure \ref{fig:symdiff_dens}) that most of these
+regions are small. Figure \ref{fig:symmetric_differences} illustrates
+three scenarios including the best, typical (closest to the median),
+and worst cases with respect to these symmetric differences. In the
+best case, we see that the buffers encompassing a very small, isolated
+pedestrian facility are practically identical. In the more typical
+case, slight differences are visible upon close inspection. Finally,
+in the worst case, which was located in a remote portion of Central
+Washingtion, it is apparent that the buffers are very different, with
+the PostGIS buffer only a subset of the larger ESRI-generated buffer.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.65\textwidth]{../results/symdiff_dens}
+  \caption{Density plot showing the distribution of symmetric
+    differences between PostGIS- and ESRI-generated sausage buffers.}
+  \label{fig:symdiff_dens}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\textwidth]{./figs/symmetric-differences}
+  \caption{Three buffers illustrating range of symmetric
+    differences. From left to right: buffer with smallest difference,
+    buffer with difference nearest the median, and buffer with largest
+    symmetric difference. PostGIS buffered regions are denoted by
+    diagonal crosshatching, while ESRI buffered regions are
+    dot-shaded.}
+  \label{fig:symmetric_differences}
+\end{figure}
+
+We believe that the primary source of differences between our PostGIS
+and ESRI generated buffers to be the result of topological differences
+and errors in the network dataset.  As discussed in the methods
+section, we imported Open Street Map data into PostGIS in such a way
+as to preserve the network topology. Thus, calculations performed with
+PostGIS / PgRouting will account for non-apparent discontinuities,
+such as where a bridge crosses over another facility in 3d space.
+
+The network used for creation of buffers in ArcGIS, on the other hand,
+was the result of constructing a network dataset from a shapefile
+exported from PostGIS. Because shapefiles do not preserve topology,
+ArcGIS is forced to infer topology. Thus, in our hypothetical
+situation where a bridge crosses over another facility, the
+reconstructed network may erroneously connect between the bridge and
+the road below. In that case, we would expect to see a smaller PostGIS
+buffer and a larger ESRI buffer, due to the greater network
+connectivity in the ESRI case. While the mean difference in area
+between our two sets of buffers was not significant in the statistical
+sense, the observed differences in our sample are consistent with the
+logic that the more connected, inferred network topology would yield
+larger buffers.
+
+Our worst case buffer pair suggests an an extreme version of this size
+bias with respect to network topology---namely that potential
+topological errors in the Open Street Maps data could be corrected
+when imported into ArcGIS via a shapefile. These two buffers are
+located in a remote portion of Central Washington, where we might
+expect network topology errors to go uncorrected, in this instance
+resulting in a false discontinuity. In such instances, the PostGIS
+buffer would again be smaller than the ArcGIS buffer. Finally, while
+most of the PostGIS-generated buffers were smaller, about 15 percent
+were larger than their ESRI-generated sibling, suggesting that
+inferred topologies may not always be more connected than the topology
+present in the OSM data.
+
+\section*{Conclusions}
+This paper set out to validate, in a fully reproducible way, the
+sausage buffer approach to creating network-based travel catchment
+geographies. We accomplished this by developing procedures using the
+PostgreSQL relational database with PostGIS and PgRouting
+extensions. We conclude that the sausage buffering method is indeed a
+valid and replicable approach to generating network-based travel
+catchments, and that PostgreSQL, PostGIS, and PgRouting are 
+suitable and offer important benefits for these purposes.
+
+The sausage buffering procedures presented here achieve a higher
+standard of reproducibility than previously published sausage
+buffering research, in that these are explicitly designed to allow for
+automated analyses that are less subject to human error. In addition
+to our procedures being fully automated through computer code, the
+free and open source nature of PostgreSQL, PostGIS, and PgRouting
+confer further benefits to reproducibility. Whereas the true
+implementation of ESRI's network analyst is not known outside of ESRI,
+all of PgRouting's distance calculations can be scrutinized
+directly. Indeed, had the service area calculations in ArcGIS been
+fully transparent in this manner, \citeauthor{Forsyth2014sausage} may
+not have had cause to develop sausage buffers in the first place.
+
+In addition to establishing a reproducible method for creating sausage
+buffers, we also conclude that the sausage buffers created via PostGIS
+are generally comparable to sausage buffers created by
+ArcGIS. Differences between sausage buffers created in the two
+programs are more likely to be due to differences in how network
+datasets are constructed, than to differences in how the buffers are
+computed.
 

### 2015-06-15 analysis/040-import-esri-buffers.sh

@@ -1,0 +1,13 @@
+#!/usr/bin/env bash
+
+DB=pgismethods
+
+shp2pgsql -d -s2926 \
+          ./data/esri_sausage_1000m/sausage_1000m_buffer_round_end.shp \
+          public.esri_1000m_round |
+    psql $DB
+
+shp2pgsql -d -s2926 \
+          ./data/esri_sausage_1000m/sausage_1000m_dissolve_flat_end.shp \
+          public.esri_1000m_flat |
+   psql $DB

### 2015-06-15 analysis/050-calc-buffer-areas.sh

@@ -1,0 +1,18 @@
+DB=pgismethods
+
+echo "COPY (SELECT ST_Area(sb_1000m.geom_poly) AS postgis,
+              ST_Area(esri_1000m_round.geom) AS esri,
+              sb_1000m.id as id
+              FROM gis.sb_1000m
+              LEFT JOIN public.esri_1000m_round
+              ON gis.sb_1000m.id = public.esri_1000m_round.id_t)
+      TO STDOUT WITH CSV HEADER;" |
+    psql $DB > ./results/buffer_areas.csv
+
+echo "COPY (SELECT ST_Area(ST_SymDifference(sb_1000m.geom_poly, esri_1000m_round.geom)) AS pg_esr,
+              sb_1000m.id as id
+              FROM gis.sb_1000m
+              LEFT JOIN public.esri_1000m_round
+              ON gis.sb_1000m.id = public.esri_1000m_round.id_t)
+      TO STDOUT WITH CSV HEADER;" |
+    psql $DB > ./results/buffer_symdiffs.csv

### 2015-06-15 analysis/055-compare-buffers.R

@@ -1,0 +1,57 @@
+library(ggplot2)
+library(reshape2)
+library(BEST)
+
+## Read our datasets
+areas_wide <- read.csv("./results/buffer_areas.csv")
+symdiffs_wide <- read.csv("./results/buffer_symdiffs.csv")
+
+## Convert to tidy/long format and ft^2 -> km^2
+ft2tokm2 <- function(x) x * 9.2903e-8  
+areas <- melt(areas_wide, value.name="sqft", variable.name="type", id.vars=c("id"))
+areas$km2 <- ft2tokm2(areas$sqft)
+symdiffs <- melt(symdiffs_wide, value.name="sqft", variable.name="type", id.vars=c("id"))
+symdiffs$km2 <- ft2tokm2(symdiffs$sqft)
+
+## Descriptive density plots of areas and symmetric differences
+area_dens <- ggplot(areas, aes(x=km2, color=type, linetype=type)) +
+  scale_color_brewer(type="qual", palette=2) +
+  geom_density() + theme_bw()
+ggsave("./results/area_dens.pdf", width=5, height=2.5)
+
+symdiff_dens <- ggplot(symdiffs, aes(x=km2)) +
+  geom_density() + theme_bw()
+ggsave("./results/symdiff_dens.pdf", width=5, height=2.5)
+
+## Statistically compare areas of pg buffers and esri round ended
+## buffers using BEST
+best_areas_pg_esr <- BESTmcmc(y1=ft2tokm2(areas_wide$postgis),
+                              y2=ft2tokm2(areas_wide$esri))
+
+## Power analysis, takes quite a while to run
+## bpwr_areas_pg_esr <- BESTpower(best_areas_pg_esr,
+##                                N1=length(areas_wide$postgis),
+##                                N2=length(areas_wide$esri_round), 
+##                                ROPEm=c(-0.0314,0.0314),
+##                                maxHDIWm=1.0, nRep=1000) 
+
+## Plots from BEST
+pdf("./results/best_areas_pg_esr_mean.pdf", width=8, height=4)
+plot(best_areas_pg_esr, "mean")
+dev.off()
+
+pdf("./results/best_areas_pg_esr_sd.pdf", width=8, height=4)
+plot(best_areas_pg_esr, "sd")
+dev.off()
+
+pdf("./results/best_areas_pg_esr_effect.pdf", width=8, height=4)
[…]

### 2015-10-03 article/article.tex
PMH changes, 20151002
PMH changes, 20151002

--- 2015-04-24 article/article.tex
@@
 \usepackage[authordate,strict,autolang=hyphen,bibencoding=inputenc,doi=true,isbn=false,annotation=false]{biblatex-chicago}
+\usepackage{graphicx}
+\DeclareGraphicsExtensions{.pdf,.png,.jpg}
+\usepackage[section]{placeins}
 \addbibresource{../references/references.bib}
+\usepackage{booktabs}
+\usepackage[affil-it]{authblk} % pretty author nameuse
+\usepackage{verbatim} % mono spaced as needed
+\usepackage{color} % for PMH's comments
+\usepackage{soul} % for strikethrough
 
-\title{Reproducible Methods for Urban Form Analysis: A Case Study of PostGIS}
-\author{Phil Hurvitz, \ldots}
+% italicize 'et al.'
+\renewbibmacro*{name:andothers}{% Based on name:andothers from biblatex.def
+  \ifboolexpr{
+    test {\ifnumequal{\value{listcount}}{\value{liststop}}}
+    and
+    test \ifmorenames
+  }
+    {\ifnumgreater{\value{liststop}}{1}
+       {\finalandcomma}
+       {}%
+     \andothersdelim\bibstring[\emph]{andothers}}
+    {}}
+
+%---------------------------------------- title, author, date
+\title{Reproducible sausage network buffers for built environment measurement: an open source GIS approach}
+% \author[*]{Philip M. Hurvitz}
+% \author[*]{}
+% \author[*]{Eric J. Howard}
+\author{Philip M. Hurvitz%
+  \thanks{Electronic address: \texttt{phurvitz@uw.edu.edu}; Corresponding author}}
+
+\author{Peter Schmiedeskamp%
+  \thanks{Electronic address: \texttt{pschmied@uw.edu}}}
+
+\author{Eric J. Howard%
+  \thanks{Electronic address: \texttt{ejhoward@uw.edu}}}
+
+\affil{Department of Urban Design and Planning, University of Washington}
+
 \date{}
 
+%----------------------------------------
 %%% BEGIN DOCUMENT
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris in
-hendrerit ligula. Cras malesuada auctor egestas. In eu placerat
-dolor. Vestibulum rhoncus risus nulla. Praesent consequat diam nec
-neque hendrerit, ac convallis turpis feugiat. Nulla quis nisi
-lectus. Suspendisse ipsum eros, facilisis at nulla sed, dignissim
-rhoncus erat. Suspendisse ut odio quis justo egestas condimentum eu
-vitae mauris. Donec tincidunt sapien ex, at porta arcu scelerisque
-in. Nunc porttitor risus et rhoncus venenatis. Sed aliquet rhoncus
-interdum. Quisque nec lacinia risus. Nulla mi nulla, varius ac elit
-iaculis, blandit lobortis augue. In pharetra erat ut ipsum pulvinar,
-id blandit purus mollis.
+\section*{Background}
+Associations between built environment exposures and behaviors or health outcomes are of great interest in public health research.
+Geographic Information Systems (GIS) methods for quantifying exposure to features within the built environment have developed rapidly, from census-unit approaches \parencite{Morland2002a}, to Euclidan (`circular' or `crow-flies') buffers around geocoded home locations \parencite{Moudon2002}, to network-distance buffers \parencite{Frank2005}.
+Buffer measures have become preferred because they capture only those features near to specific locations and avoid the Modifiable Areal Unit Problem \parencite{Openshaw1984}, with network buffers favored due to their ability to mask out areas that are reasonably inaccessible by foot or vehicle \parencite{Oliver2007}.
 
-\cite{Forsyth2014sausage}
+\textcite{Forsyth2014sausage}
+identified a threat to replicability for studies using such
+network-based buffers, due to changes in how Esri's
+market-leading GIS software has calculated them across different software versions. In response,
+the authors proposed ``sausage'' network buffers,
+created via a transparent algorithm, for use in measuring aspects of
+urban form related to food and physical activity.
+The algorithm was operationalized using Esri's ArcGIS desktop GIS software on Microsoft Windows, but owing to its simplicity and transparency, was described as potentially
+replicable and cross-platform. Their paper called on others to validate sausage buffers
+created using other GIS software on other platforms.
+
+This paper serves several purposes.
+It answers \citeauthor{Forsyth2014sausage}'s call by presenting and validating an alternate method for the creation of sausage buffers using readily available, cross-platform, free and open source software; in doing so it also introduces and demonstrates the specific software used for development of the method, PostGIS \parencite{ThePostGISDevelopmentGroup2008}, as an alternative to proprietary, commercial GIS software.
+PostGIS is an extension to the free and open source relational database PostgreSQL \parencite{ThePostgreSQLGlobalDevelopmentGroup2008}.
+In this paper we focus on the use of network and buffer analysis, but PostGIS contains a large set of geoprocessing functions common to most desktop GIS applications, such as geometric overlay, coordinate projections, and raster analysis.
+
+%Let's mention something about reproducibility up front.
+The methods presented here are also in part a response to a growing
+call from numerous disciplines to ensure computer aided
+analyses are truly reproducible. The basic standard for reproducibility is
+that the data used are made available, and that analyses are shared in the
+form of computer code \parencite{Peng2011computational}. Notably, the
+methods presented here are reproducible because the analyses are fully
+automated using computer code, rather than being reliant on a set of
+manual procedures. The code and the data have been made publicly
+available, so that other researchers may both review and benefit from this work.\footnote{need to package up code and data, upload to UW's
+  repository, and then see about generating a DOI.}  Further, building
+these methods atop free and open source software achieves a higher
+standard of reproducibility by eliminating dependencies on any ``black
+box'' methods within proprietary software that may be necessary in data processing. In principle, this means that every piece of
+software, as well as the specific analytic code and data used in these methods, is open to scrutiny.
+
+% Relevance to planning and epidemiology
+Developing geoprocessing methods around free and open source tools
+holds particular benefits for disciplines such as epidemiology, public
+policy, and urban planning because of the reduced barriers of adoption
+for practitioners and advocates outside of larger institutions. Academic
+institutions and larger public agencies may be able to afford costly software
+applications. However, smaller agencies or advocacy groups wishing to
+perform quantitative analysis of built environment might be unable to
+make use of even transparently described analyses that are reliant on proprietary software, simply due to the
+cost of software licensing fees.
+
+% Not a literature review, per se. Since we are developing a methods
+% paper largely in response to one particularly important method, I'm
+% not sure how important a full-blown review is. Open to suggestions,
+% however. This could potentially be split off and developed into
+% something more complete. 
+% PMH: I agree, we aren't doing an evaluation of the use of PostGIS in research.
+Finally, the PostGIS platform is especially beneficial in that it
+naturally facilitates reproducible, scripted analyses, as well as 
+provides the core infrastructure for data storage and management on even very large research
+projects. \textcite{hurvitz2014emerging} describe the use of
+PostgreSQL/PostGIS to manage and process geospatial and timeseries
+data from a variety of sources including Global Positioning Systems (GPS) and
+accelerometers for very large data sets. However, we are neither first nor alone in identifying
+PostGIS for use in research. A keyword search for ``PostGIS'' using
+ISI Web of Science identified 35 \textcolor{red}{[I got 50 today!]} articles across a number of
+disciplines describing the use of PostGIS in research. In the
+transportation sector, \textcite{Wang2015routable} also used PostGIS
+to store and analyze a large number of GPS traces in order to infer
+characteristics about roadways such as position and rules, for the
+purposes of generating routable roadway networks. In another recent
+article, \textcite{Brovelli2015FOSS} describe the use of PostGIS
+together with GeoServer to collect data about roadway pavement
+conditions from the general public. Notable in this study is the
+potential for PostGIS in supporting digitally-mediated public
+participation.
+However, a PubMed search for ``PostGIS'' resulted in only one paper \parencite{Rowlingson2013}.
+In contrast, Web of Science and PubMed searches for ``ArcGIS'' resulted in 1,674 and 394 citations, respectively.
+This indicates that a number of researchers who could benefit from the use of PostGIS may simply be unaware of its existence.
+
+\section*{Methods}
+% Set-up and software pgRouting 2.x on OpenBSD 5.7 i5
+Following the process outlined in 
+\textcite{Forsyth2014sausage} and described in step-by-step detail in \textcite{ Forsyth2012proto}, we implemented the
+sausage buffering technique using PostgreSQL 
+with PostGIS and the pgRouting \parencite{pgRoutingContributors2013} extension, which implements a number of network routing algorithms, including Dijkstra's shortest path \parencite{Dijkstra1959}, which is also using in the ArcGIS Network Analyst exension.
+Using that technique we
+generated 200 network buffers around randomly selected points. We then
+compared those results to buffers analogously created in ArcGIS Desktop
+10.2.X, for use as a basis for comparison with the results presented
+by \citeauthor{Forsyth2014sausage}. Table \ref{tab:software} lists
+software versions used in this analysis.
+
+\begin{table}[h!]
+  \centering
+  \caption{Software used to generate PostGIS and ArcGIS sausage buffers}
+  \begin{tabular}{llr}
+  \toprule
+  Buffer type & Software & Version \\
+  \midrule
+PostGIS/pgRouting & OpenBSD (Operating System) & 5.7 \\
+              & PostgreSQL & 9.4.2 \\
+              & PostGIS Extensions & 2.1.7 \\
+              & ggRouting & 2.0.0 \\
+              & Osm2pgrouting & 2.0.0 \\
+              &  &  \\
+  ArcGIS & Microsoft Windows & Eric? \\
+              & ArcGIS? & 10.2.X Eric? \\
+  \bottomrule
+\end{tabular}
+\label{tab:software}
+\end{table}
+
+In overview, the process for creating a sausage buffer entails three
+key steps. First, a point of origin is identified on a network of
+transportation facilities. Next, a distance threshold is selected. The
+network is traversed outward along all possible paths from the point
+of origin until reaching the distance threshold. The resulting subset
+reflects portions of the network reachable within the chosen
+distance. Finally, the segments of the network subset are buffered
+according to some chosen Euclidean buffer radius, resulting in a set of polygons
+reminiscent of sausages. These are ``dissolved'' into a single polygon,
+forming the final buffer. Figure \ref{fig:sausage-steps} illustrates
+these steps.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\textwidth]{./figs/sausage-steps}
+  \caption{The three primary steps in creating a
+    sausage buffer. Left: an origin is selected. Center: the portions
+    of the network reachable within a given distance threshold (here,
+    1km) are identified. Right: lines are buffered and dissolved.
+\textcolor{red}{LET'S MAKE THE POINT OF ORIGIN, NORTH ARROW, AND SCALE BAR MORE PROMINENT}
+}
+	\label{fig:sausage-steps}
+\end{figure}
+
+%\subsection*{Data sources and network import}
+To create the routable network required for generating the buffers, we
+imported the Washington State portion of the Open Street Map (OSM) (\copyright\ OpenStreetMap contributors)
+into PostGIS / pgRouting using the open source osm2pgrouting
+utility. \textcolor{red}{Did we use GeoFabrik? If so we should describe.} Because much of our work deals with pedestrian travel, we
+excluded highway facilities, which often do not allow for pedestrian
+access. The osm2pgrouting import strategy yielded two primary
+benefits. First, osm2pgrouting facilitates reproducible data
+importation through the use of an XML-based configuration file, which 
+allows facility exclusions to be carried out consistently, such as our exclusion of highways as non-traversible for
+pedestrians. Secondly, osm2pgrouting
+preserves the topology of the OSM network, rather than relying on
+topology from the data set's simple vector geometry. Maintaining network topology is important for enforcing physical constraints such as over/underpasses or bridges.
+
+%\subsection*{Generation of buffers}
+
+% Creation of buffer code in database
+% Random sample of 200 vertices as points to construct buffers around
+% Define sausage buffering procedure in PLSQL
+% Construct 1000m buffer (net dist), 100m (arc dist)
+% Rounded versus squared ends
+We implemented the sausage buffering algorithm in PL/PgSQL, which provides basic SQL functionality as well as procedural characteristics such as control structures and complex computations, with functions saved as stored procedures within the database.
+We created two functions; the first generates a
+sausage buffer around a single point of origin, and the second uses
+the first function to generate sausage buffers around an arbitrary
+number of points.
+
+Because a point of origin may not be located precisely on the network,
+the sausage buffering function begins by identifying and `snapping' to the
+vertex in the network nearest to the identified point of origin. To limit the
+required number of distance calculations required, the function then
+identifies vertices within a circular buffer with radius equal to the
+chosen network distance threshold, which represents the maximum
+possible distance one could travel on the network in any
+direction. We then generate the route to each vertex within this
+circular buffer. Because the network distance from the origin to any given vertex could exceed the threshold, we truncated each route using the threshold distance.\footnote{We
+  believe this approach to be computationally inefficient, however
+  necessary due to the lack of a built-in service area function in
+  pgRouting. Our hope is that future versions of pgRouting will
+  include service area analysis, thus simplifying and speeding up
+  creation of these buffers.} The resulting set of lines are then
+buffered according to the buffer radius, and the resulting polygons
+are dissolved into a single geometry.
+
+The PL/PgSQL implementation was used to generate buffers originating from 200 randomly
+selected points, which were selected from the road data set's network vertices to ensure that they were not located in
+roadless areas of the state. Then, in order to compare buffers
+created using the PostGIS implementation with ArcGIS-generated buffers, we exported the origin points and OSM network data as shapefiles and followed the protocols described in
+\textcite{Forsyth2012proto}, but using rounded buffer ends as described in \textcite{Forsyth2014sausage}.
+
+Finally, we made two comparisons between the two sets of
+buffers. First to assess whether differences in
+the planimetric area between these two groups were statistically significant, we
+used a Bayesian procedure somewhat analogous to a t-test called BEST
+\autocite{Kruschke2014,Kruschke2013}. \textcolor{red}{Why this method instead of a t-test?} Second, to assess the possibility of
+spatial misalignment between similarly sized buffers, we computed the
+symmetric difference of the two sets of sausage buffers and then
+calculated the area of the non-overlapping regions.
+
+\section*{Results}
+Overall, the differences between the PostGIS and ArcGIS generated buffers were quite low. Figure
+\ref{fig:area_dens} shows the distribution of areas of these two sets of
+buffers. Both exhibit the same somewhat bimodal shape, with greater
+differences observed among smaller buffers.
+Although the PostGIS-generated buffers were slightly smaller, the mean difference's 95\% confidence interval was -0.23 to 0.04, indicating no significant difference in size using the two methods.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.65\textwidth]{../results/area_dens}
+  \caption{Density plot showing the distributions PostGIS- and
+    ArcGIS-generated sausage buffers across the range of sizes.}
+  \label{fig:area_dens}
+\end{figure}
+
+The symmetric differences between the two sets of
+buffers are small (figure \ref{fig:symdiff_dens}), \textcolor{red}{with X being less than Y}.
+\textcolor{cyan}{Can we add some numbers about the relative size of the differences to the size of the buffers?}.
+Figure \ref{fig:symmetric_differences} illustrates
+three scenarios including the best, typical (closest to the median),
+and worst cases with respect to these symmetric differences. In the
+best case, the buffers encompassing a very small, isolated
+pedestrian facility are practically identical. In the more typical
+case, slight differences are visible upon close inspection. Finally,
+in the worst case, which was located in a remote portion of Central
+Washingtion, the buffers are very different, with
+the PostGIS buffer only a subset of the larger ArcGIS-generated buffer.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.65\textwidth]{../results/symdiff_dens}
+  \caption{Density plot showing the distribution of symmetric
+    differences between PostGIS- and ArcGIS-generated sausage buffers.}
+  \label{fig:symdiff_dens}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=\textwidth]{./figs/symmetric-differences}
+  \caption{Three buffers illustrating a range of symmetric
+    differences. Left: buffer with smallest difference. Center:
+    buffer with difference nearest the median. Right: buffer with largest
+    symmetric difference. PostGIS buffered regions are denoted by
+    diagonal crosshatching, while ArcGIS buffered regions are
+    dot-shaded. \textcolor{red}{Separate these images from each other with whitespace. Subfig or some such. Also can we need a better polygon fill that shows up at a readable size/scale.}}
+  \label{fig:symmetric_differences}
+\end{figure}
+
+\section*{Discussion}
+This study presented a new method for generating sausage buffers using free and open source GIS software, and compared the results of the new method to those of an existing method that relies on the leading proprietary GIS application, ArcGIS.
+
+There were measurable, but not signiticant differences between buffers generated using the two methods. The primary source of differences between the PostGIS
+and ArcGIS generated buffers is likely the result of topological differences
+and errors in the network dataset. Osm2pgrouting imports OSM data into PostGIS while preserving network topology. Calculations performed with
+PostGIS / pgRouting will properly handle non-apparent discontinuities,
+such as the aforementioned cases of over/underpasses. The network used for creation of buffers in ArcGIS, on the other hand,
+was exported as a shapefile from PostGIS. Because shapefiles do not preserve topology,
+ArcGIS is forced to infer topology from simple ``spaghetti'' geometry. Thus, in the 
+situation where a bridge crosses over another roadway, the
+reconstructed network may erroneously connect between the bridge and
+the road below. \textcolor{red}{Can we verify this?} In that case, the PostGIS
+buffer would be smaller than the ArcGIS buffer, due to the greater, but erroneous 
+connectivity in the ArcGIS network. While the mean difference in area
+between our two sets of buffers was not statisitcally significant, the observed differences in our sample are consistent with the
+logic that the more connected, inferred network topology would yield
+larger buffers.
+\textcolor{red}{OK, sorry to ask, but we really should export back from shapefile to PostGIS as spaghetti for at least one of the bad cases of difference to test whether that was indeed the case.
+Or examine the roads in that worst case to verify that was happening.
+Then we needn't rely on a logical argument but can point directly to the cause.}
+
+\textcolor{red}{I don't think we need this.}\st{Our worst case buffer pair suggests an an extreme version of this size
+bias with respect to network topology---namely that potential
+topological errors in the Open Street Maps data could be corrected
+when imported into ArcGIS via a shapefile. These two buffers are
+located in a remote portion of Central Washington, where we might
+expect network topology errors to go uncorrected, in this instance
+resulting in a false discontinuity. In such instances, the PostGIS
+buffer would again be smaller than the ArcGIS buffer.} Finally, while
+most of the PostGIS-generated buffers were smaller, about 15 percent
+were larger than their ArcGIS-generated cognate, suggesting that
+inferred topologies may not always be more connected than the topology
+present in the OSM data.
+
+\section*{Conclusions}
+This paper set out to validate, in a fully reproducible way, the
+sausage buffer approach to creating network-based travel catchment
+areas. We accomplished this by developing procedures using the
+PostgreSQL relational database and its PostGIS and pgRouting
+extensions.
+
+In addition to establishing a reproducible method for creating sausage
+buffers, the buffers created using PostGIS and ArcGIS 
+are generally comparable. Differences between sausage buffers created in the two
+programs are more likely to be due to differences in how network
+datasets are constructed, than to differences in the underpinnings of how the buffers are
+computed.
+
+PostgreSQL, PostGIS, and pgRouting provide
+suitable analytics and offer important benefits for several purposes.
+With automated analyses that are less subject to human error, the sausage buffering procedures presented here achieve a higher
+standard of reproducibility than the previously published research. In addition
+to our procedures being fully automated through computer code, the
+free and open source nature of PostgreSQL, PostGIS, and pgRouting
+confer further benefits to reproducibility. Whereas the underlying algorithms used in ArcGIS's Network Analyst are not known outside of Esri,
+all of pgRouting's distance calculations and logic can be scrutinized
+directly in the source code. Indeed, had the service area calculations in ArcGIS been
+fully transparent in this manner, \citeauthor{Forsyth2014sausage} may
+not have had cause to develop sausage buffers in the first place.
 

### 2015-10-03 article/article.tex
PMH 20151002, again
PMH 20151002, again

@@
 holds particular benefits for disciplines such as epidemiology, public
-policy, and urban planning because of the reduced barriers of adoption
+policy, and urban planning because of the reduced barriers to adoption
 for practitioners and advocates outside of larger institutions. Academic
@@
 data from a variety of sources including Global Positioning Systems (GPS) and
-accelerometers for very large data sets. However, we are neither first nor alone in identifying
+accelerometers for very large data sets. We are admittedly neither first nor alone in identifying
 PostGIS for use in research. A keyword search for ``PostGIS'' using
@@
 together with GeoServer to collect data about roadway pavement
-conditions from the general public. Notable in this study is the
+conditions from the general public. Notable in that study is the
 potential for PostGIS in supporting digitally-mediated public
@@
 In contrast, Web of Science and PubMed searches for ``ArcGIS'' resulted in 1,674 and 394 citations, respectively.
-This indicates that a number of researchers who could benefit from the use of PostGIS may simply be unaware of its existence.
+This indicates that a number of researchers who could potentially benefit from the use of PostGIS may simply be unaware of its existence.
 
@@
 sausage buffering technique using PostgreSQL 
-with PostGIS and the pgRouting \parencite{pgRoutingContributors2013} extension, which implements a number of network routing algorithms, including Dijkstra's shortest path \parencite{Dijkstra1959}, which is also using in the ArcGIS Network Analyst exension.
+with PostGIS and the pgRouting \parencite{pgRoutingContributors2013} extension, which implements a number of network routing algorithms, including Dijkstra's shortest path \parencite{Dijkstra1959}, which is also used in the ArcGIS Network Analyst exension.
 Using that technique we
-generated 200 network buffers around randomly selected points. We then
+generated 200 sausage buffers around randomly selected points. We then
 compared those results to buffers analogously created in ArcGIS Desktop
@@
 into PostGIS / pgRouting using the open source osm2pgrouting
-utility. \textcolor{red}{Did we use GeoFabrik? If so we should describe.} Because much of our work deals with pedestrian travel, we
+utility. \textcolor{red}{Did we use GeoFabrik? If so we should describe.} Because much of the work in measuring built environment exposures deals with pedestrian travel, we
 excluded highway facilities, which often do not allow for pedestrian
@@
 preserves the topology of the OSM network, rather than relying on
-topology from the data set's simple vector geometry. Maintaining network topology is important for enforcing physical constraints such as over/underpasses or bridges.
+topology created from the data set's simple vector geometry. Maintaining network topology is important for enforcing physical constraints to traversability such as over/underpasses or bridges.
 
@@
 
-The PL/PgSQL implementation was used to generate buffers originating from 200 randomly
-selected points, which were selected from the road data set's network vertices to ensure that they were not located in
+The PL/PgSQL implementation was used to generate buffers originating from 200 points, which were randomly selected from the road data set's network vertices to ensure that they were not located in
 roadless areas of the state. Then, in order to compare buffers
@@
 \ref{fig:area_dens} shows the distribution of areas of these two sets of
-buffers. Both exhibit the same somewhat bimodal shape, with greater
+buffers. Both exhibit the same bimodal shape, with greater
 differences observed among smaller buffers.
@@
 The symmetric differences between the two sets of
-buffers are small (figure \ref{fig:symdiff_dens}), \textcolor{red}{with X being less than Y}.
+buffers were small (figure \ref{fig:symdiff_dens}), \textcolor{red}{with X being less than Y}.
 \textcolor{cyan}{Can we add some numbers about the relative size of the differences to the size of the buffers?}.
@@
 Washingtion, the buffers are very different, with
-the PostGIS buffer only a subset of the larger ArcGIS-generated buffer.
+the PostGIS buffer being only a subset of the larger ArcGIS-generated buffer.
 
@@
 \section*{Discussion}
-This study presented a new method for generating sausage buffers using free and open source GIS software, and compared the results of the new method to those of an existing method that relies on the leading proprietary GIS application, ArcGIS.
+This study presented a new method for generating sausage buffers using free and open source GIS software, and compared the results of this method to those of an existing method that relies on the leading proprietary GIS application, ArcGIS.
 
 There were measurable, but not signiticant differences between buffers generated using the two methods. The primary source of differences between the PostGIS
-and ArcGIS generated buffers is likely the result of topological differences
+and ArcGIS generated buffers is \textcolor{red}{likely} the result of topological differences
 and errors in the network dataset. Osm2pgrouting imports OSM data into PostGIS while preserving network topology. Calculations performed with
-PostGIS / pgRouting will properly handle non-apparent discontinuities,
+PostGIS / pgRouting will properly handle not readily apparent discontinuities,
 such as the aforementioned cases of over/underpasses. The network used for creation of buffers in ArcGIS, on the other hand,
@@
 are generally comparable. Differences between sausage buffers created in the two
-programs are more likely to be due to differences in how network
+programs are more likely due to differences in how network
 datasets are constructed, than to differences in the underpinnings of how the buffers are
